---
title: "Final Project Interterm"
date: "10/7/2020"
output:
  html_document:
    highlight: zenburn
    theme: flatly
    toc: yes
    toc_float: yes
  toc: default
authors: Conor McGrath, Tina Wang, Tiffany Tseng, Michael Mayor, Brock Gallagher
---
# Introduction

Team Name: L1
Team Members: Conor McGrath, Tina Wang, Tiffany Tseng, Michael Mayor, Brock Gallagher

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Data sets

```{r}
library(tidyverse)
Baseball2019 <- read.csv("./stats2019.csv")
Baseball2020 <- read.csv("./stats2020.csv")
```


# Combine names columns for each table. 

```{r}

Baseball2019$Name <- paste(Baseball2019$first_name, Baseball2019$last_name, sep = ' ')

Baseball2020$Name <- paste(Baseball2020$first_name, Baseball2020$last_name, sep = ' ')
```

# Create average HR per AB column

```{r}
Baseball2019$AvgHrPerAB <- round(Baseball2019$b_home_run / Baseball2019$b_ab, 2)

Baseball2020$AvgHrPerAB <- round(Baseball2020$b_home_run / Baseball2020$b_ab, 2)
```

# Select necesarry columns

```{r}
Baseball2019 <- Baseball2019 %>%
  select(Name, AvgHrPerAB, exit_velocity_avg, launch_angle_avg)

Baseball2020 <- Baseball2020 %>%
  select(Name, AvgHrPerAB, exit_velocity_avg, launch_angle_avg)
```

# Innerjoin on new player name column

```{r}
BaseballFinal <- inner_join(Baseball2019, Baseball2020, by = "Name")

BaseballFinal <- BaseballFinal %>%
  select(1, 2, 5, 3, 6, 4, 7)

colnames(BaseballFinal) <- c("Name", "AvgHrPerAB2019", "AvgHrPerAB2020", "EV2019", "EV2020", "LA2019", "LA2020")

```


# Create column with difference in EV and LA from 2019 to 2020

```{r}
BaseballFinal$EVDiff <- round(BaseballFinal$EV2020 - BaseballFinal$EV2019, 2)

BaseballFinal$LADiff <- round(BaseballFinal$LA2020 - BaseballFinal$LA2019, 2)

BaseballFinal$HRperABDiff <- round(BaseballFinal$AvgHrPerAB2020 - BaseballFinal$AvgHrPerAB2019, 2)
```

# Select only relevant columns

```{r}
BaseballFinal <- BaseballFinal %>%
  select(1,8,9,10)
```

# Created a binary variable for Avg HR Diff
```{r}
BaseballFinal$AvgHR <- if_else(BaseballFinal$HRperABDiff < 0, "Decreased", "Increased")

BaseballFinal$AvgHR <- as.factor(BaseballFinal$AvgHR)
```


# Created 3 scatterplotts

```{r}

library(plotly)

ggplot(BaseballFinal, aes(x=EVDiff, y=HRperABDiff)) + geom_point(aes(color=AvgHR)) + geom_smooth() + scale_color_manual(values = c("red", "green")) + theme_minimal()

ggplot(BaseballFinal, aes(x=LADiff, y=HRperABDiff)) + geom_point(aes(color=AvgHR)) + geom_smooth() + scale_color_manual(values = c("red", "green")) + theme_minimal()

ggplotBB3 <- ggplot(BaseballFinal, aes(text = paste("Player Name:", Name, "<br>Change in HR per AB ratio:", HRperABDiff, "<br>Change in Avg EV:", EVDiff, "<br>Change in Avg LA:", LADiff, "<br>Avg HR:", AvgHR), x=EVDiff, y=LADiff, color = AvgHR)) + geom_point() + scale_color_manual(values = c("red", "green")) + ggtitle("Effects of Launch Angle and Exit Velocity", subtitle = "Source: Baseball Savant") + theme_minimal()

ggplotly(ggplotBB3, tooltip = "text")



```

